---
title: "Covid91 vaccine study Final2022"
author: "MarenaO"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 5
  pdf_document:
    toc: yes
    toc_depth: '5'
---


```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final_Data.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
males = filter(FinalData, sex =="M")
females = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```


# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

```{r}
dd2 <- males %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


This plot represents the population of males and their relationship to getting covid-19 by taking either the vaccine or the placebo. The plot shows that around 68% of males who got covid took the placebo. Whereas about 32% of males who got covid took the vaccine. Therefore, more males who got covid took the placebo, whereas the males who got covid from taking the vaccine were not as likely to contract covid as much. This means that the vaccine seems to be effective for the males in this dataset. 
  There were two plots that came out. The first plot shows two bars for the treatment, either the placebo or the vaccine. Along the x-axis was the amount of people there was, and along the y-axis shows the infected from covid, either placebo or vaccine. It shows the number of people infected from covid or from the placebo. The second plot also shows two bars for the treatments, either the placebo or the vaccine. It shows the percentage of people infected from covid or from the placebo.



### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=males)
rowPerc(table1)
colPerc(table1)
```

The numerical results for the males shows the percentages of males who got covid while taking the vaccine and males who got covid while taking the placebo. It shows that around 67% of the males who took the placebo got covid. Whereas only 32% of males who took the vaccine got covid. This comes to show that the vaccine is effective towards males. The odds of getting covid-19 while taking the vaccine is 2:1.

### Inferential 



```{r}
chisq.test(table1)
chisqtestGC(table1)
malesfish <- fisher.test(table1)
malesfish
```

A Fisher’s Exact test was run and a p-value was given. The p-value came out to be 2.645e-11, which is quite a small p-value. This means that these results are not likely to occur by chance and that there is a statistically significant result.
	The odds ratio above came out to be 2.09, meaning that the males that took the vaccine were less likely to contract covid than those who took the placebo.



```{r}
effect <- malesfish$p.value/(malesfish$p.value +1)
effectivenessguy = malesfish$estimate/(malesfish$estimate +1)*100
print(effectivenessguy)
```




```{r}
low=(malesfish$conf.int[1]/(malesfish$conf.int[1] +1))*100
hi=(malesfish$conf.int[2]/(malesfish$conf.int[2] +1))*100
options(digits=2)
```


## Effectiveness Confidence Interval `r low`  to `r hi`

The odds ratio essentially shows how effective the vaccine is. The odds ratio for the males is around 67.7, with an effectiveness confidence interval of 62.51-72.51%, meaning that the vaccine is effective in protecting males from getting covid-19.


## Females

```{r}
dd2 <- females %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot represents the population of females and their relationship to getting covid-19 by taking either the vaccine or the placebo. The plot shows that around 58% of females who took the placebo got covid. Whereas about 41% of females who took the vaccine got covid.  Therefore, more females who took the placebo got covid, whereas the males who took the vaccine did not contract covid as much. This means that the vaccine seems to be effective for females in this dataset. Although, it seems to be more effective for males than females, since more females contracted covid while taking the vaccine than the males.
	There were two plots that came out. The first plot shows two bars for the treatment, either the placebo or the vaccine. Along the x-axis was the amount of people there was, and along the y-axis shows the infected from covid, either placebo or vaccine. It shows the number of people infected from covid or from the placebo. The second plot also shows two bars for the treatments, either the placebo or the vaccine. It shows the percentage of people infected from covid or from the placebo.



### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=females)
rowPerc(table1)
colPerc(table1)
```

The numerical results for the females shows the percentages of females who got covid while taking the vaccine and males who got covid while taking the placebo. It shows that around 59% of the females who took the placebo got covid. Whereas only 41% of females who took the vaccine got covid. This comes to show that the vaccine is effective towards females. The odds of getting covid-19 while taking the vaccine is 2:1.

### Inferential 



```{r}
chisq.test(table1)
chisqtestGC(table1)
femalesfish <- fisher.test(table1)
femalesfish
```

A Fisher’s Exact test was run and a p-value was given. The p-value came out to be 9e-04, which is also quite a small p-value. This means that these results are not likely to occur by chance and that there is a statistically significant result.
	The odds ratio above came out to be 1.4, meaning that the females that took the vaccine were less likely to contract covid than those who took the placebo.



```{r}
effect <- femalesfish$p.value/(femalesfish$p.value +1)
effectivenessfemales = femalesfish$estimate/(femalesfish$estimate +1)*100
print(effectivenessfemales)
```



```{r}
low=(femalesfish$conf.int[1]/(femalesfish$conf.int[1] +1))*100
hi=(femalesfish$conf.int[2]/(femalesfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

Shown above is the effectiveness for the covid-19 vaccine for females. If the effectiveness is above 50%, then the vaccine is shown to be effective. Based on the effectiveness confidence interval of 53.5%-64.01%, I can conclude that the vaccine is effective in protecting females from getting covid-19.
	The odds ratio essentially shows how effective the vaccine is. The odds ratio for the females is 59 with an effectiveness confidence interval of 53.5-64.01%, meaning that the vaccine is effective in protecting females from getting covid-19.



## LGBTQ

```{r}
dd2 <- gay %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


This plot represents the population of gays and their relationship to getting covid-19 by taking either the vaccine or the placebo. The plot shows that around13% of gays who took the placebo got covid. Whereas about 85% of gays who took the vaccine got covid. Therefore, more gays who took the vaccine got covid, whereas the gays who took the placebo did not contract covid as much. This means that the vaccine does not seem to be effective for gays in this dataset. 
There were two plots that came out. The first plot shows two bars for the treatment, either the placebo or the vaccine. Along the x-axis was the amount of people there was, and along the y-axis shows the infected from covid, either placebo or vaccine. It shows the number of people infected from covid or from the placebo. The second plot also shows two bars for the treatments, either the placebo or the vaccine. It shows the percentage of people infected from covid or from the placebo.



### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```

The numerical results for the gays shows the percentages of gays who got covid while taking the vaccine and gays who got covid while taking the placebo. It shows that 13% of the gays who took the placebo got covid. Whereas 87% of the gays who took the vaccine got covid. This comes to show that the vaccine is ineffective towards the gays at preventing the contraction of covid-19.. The odds of getting covid-19 while taking the vaccine is 0.5:3. The gays that took the vaccine were six times more likely to get covid than those who took the placebo.


### Inferential 



```{r}
chisq.test(table1)
chisqtestGC(table1)
gayfish <- fisher.test(table1)
gayfish
```


A Fisher’s Exact test was run and a p-value was given. The p-value came out to be 2e-09, which is a small p-value. This means that these results are not likely to occur by chance and that there is a statistically significant result.
The null hypothesis is rejected in favor of the alternative hypothesis. The vaccine is ineffective in preventing the likelihood of getting covid in individuals who identify as gay.
	The odds ratio above came out to be 0.15, meaning that those gays that took the vaccine were very likely to contract covid than those who took the placebo.



```{r}
effect <- gayfish$p.value/(gayfish$p.value +1)
effectivenessgay = gayfish$estimate/(gayfish$estimate +1)*100
print(effectivenessgay)
```



```{r}
low=(gayfish$conf.int[1]/(gayfish$conf.int[1] +1))*100
hi=(gayfish$conf.int[2]/(gayfish$conf.int[2] +1))*100
options(digits=2)
```


## Effectiveness Confidence Interval `r low`  to `r hi`


Shown above is the effectiveness for the covid-19 vaccine for the gays. If the effectiveness is above 50%, then the vaccine is shown to be effective. Based on the effectiveness confidence interval of 5.65%-23.78%, I can conclude that the vaccine is ineffective in protecting the gays from getting covid-19.
	The odds ratio essentially shows how effective the vaccine is. The odds ratio for the gays is 13, with an effectiveness confidence interval of 5.65-23.78%, meaning that the vaccine is not effective in protecting the gays from getting covid-19.


## Druggies

```{r}
dd2 <- Druguser %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot represents the population of drug users and their relationship to getting covid-19 by taking either the vaccine or the placebo. The plot shows that 13% of the drug users who took the placebo got covid. Whereas 87% of the drug users who took the vaccine got covid. Therefore, more gays who took the vaccine got covid, whereas the gays who took the placebo did not contract covid as much. This means that the vaccine does not seem to be effective for the drug users in this dataset. 
	There were two plots that came out. The first plot shows two bars for the treatment, either the placebo or the vaccine. Along the x-axis was the amount of people there was, and along the y-axis shows the infected from covid, either placebo or vaccine. It shows the number of people infected from covid or from the placebo. The second plot also shows two bars for the treatments, either the placebo or the vaccine. It shows the percentage of people infected from covid or from the placebo.


### Numerical

```{r}
table1 <- xtabs(~infected + treatment, data=Druguser)
rowPerc(table1)
colPerc(table1)
```

The numerical results for the drug users shows the percentages of drug users who got covid while taking the vaccine and drug users who got covid while taking the placebo. It shows that 13% of the drug users who took the placebo got covid. Whereas 87% of the drug users who took the vaccine got covid. This comes to show that the vaccine is ineffective towards the drug users at preventing the contraction of covid-19. The odds of getting covid-19 while taking the vaccine is 0.6:3.3.

### Inferential 


```{r}
chisq.test(table1)
chisqtestGC(table1)
Druguserfish <- fisher.test(table1)
Druguserfish
```

```{r}
effect <- Druguserfish$p.value/(Druguserfish$p.value +1)
effectivenessDruguser = Druguserfish$estimate/(Druguserfish$estimate +1)*100
print(effectivenessDruguser)
```

A Fisher’s Exact test was run and a p-value was given. The p-value came out to be 2e-09, which is a small p-value. This means that these results are not likely to occur by chance and that there is a statistically significant result.
The null hypothesis is rejected in favor of the alternative hypothesis. The vaccine is ineffective in preventing the likelihood of getting covid in individuals who identify as gay.
	The odds ratio above came out to be 0.15, meaning that those gays that took the vaccine were very likely to contract covid than those who took the placebo.



```{r}
low=(Druguserfish$conf.int[1]/(Druguserfish$conf.int[1] +1))*100
hi=(Druguserfish$conf.int[2]/(Druguserfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

Shown above is the effectiveness for the covid-19 vaccine for the gays. If the effectiveness is above 50%, then the vaccine is shown to be effective. Based on the effectiveness confidence interval of 5.65%-23.78%, I can conclude that the vaccine is ineffective in protecting the gays from getting covid-19.
	The odds ratio essentially shows how effective the vaccine is. The odds ratio for the gays is 13, with an effectiveness confidence interval of 5.65-23.78%, meaning that the vaccine is not effective in protecting the gays from getting covid-19.


# Conclusion


The covid-19 vaccine was not effective in protecting both the gays and the drug users from contracting covid-19. The covid-19 effectiveness was 13 for both the gays and the drug users. Although, the covid-19 vaccine was effective in protecting the males and females against covid-19. The covid-19 effectiveness for the males was around 67.7, and for the females it was 59. With this vaccine, both the females and males were protected, with the males being protected the most.
If you identify as a gay or are a drug user, then this vaccine is not recommended because it can cause the contraction of covid-19. Although, if you are a male or a female, the vaccine is shown to be effective towards preventing the contraction of covid-19. 
The null hypothesis that there is no difference in taking the vaccine can be accepted to the categories that include the gays and the drugusers. The alternative hypothesis, that the vaccine works and is effective can be accepted to the categories that include the males and the females.

